# Topological Data Analysis

------------------------------------------------------------------------

#### Dependencies

*Install the following libraries before running the pipeline.*

```{r}
install.packages(c("TDA", "ggplot2"))
```

------------------------------------------------------------------------

## Functions and paths

*Pull generalized functions for processing the discretization files and indicate paths.*

```{r}
# get helper functions
source("utilities.R")

# specify location of discretized CSVs with cell values and path for TDA outputs
csv_path <- "~/Research/data"
out_path <- "outputs"

# image (width) resolution for outputted figures
resolution <- 1500    # in pixels

# make sure output directory exists
make_dir(out_path)
```

## Persistence diagrams (PDs)

*This section will generate persistence diagrams (1D homology) for each discretized CSV provided.*

#### Generating persistence diagrams

```{r}
# indicate cell types for only including specific cells (put NA for all cells)
types <- NA   # either a single value or vector
index <- 16    # default index for cell type is last column

# recursively iterates through specified path to find CSVs and make diagrams
make_PDs(csv_path, out_path, types, index=index)
```

#### Plotting persistence diagrams

```{r}
# diagram point color and potentially specific custom plot max
color <- "orange"
# plot_max <- 25

# generate plots for PDs (this will read "maxes.rds" for the plot limits)
plot_PDs(out_path, color, resolution)
```

## Persistence landscapes (PLs)

*This section will generate persistence landscapes based on the previous section.*

#### Generating persistence landscapes

```{r}
# recursively iterates through all diagram files to make PLs
make_PLs(out_path, n=1000, disc_step=0.1)    # the discretization step can be overridden
```

#### Plotting persistence landscapes

```{r}
# specify the number of landscape functions to plot
n <- 50    # this will plot the first "n" functions

# generate plots for PLs (this will read "maxes.rds" for the plot limits) 
plot_PLs(out_path, n, resolution)
```

------------------------------------------------------------------------

## Comparing homology between groups

*Pull generalized functions for comparing groups.*

```{r}
# get helper functions
source("utilities.R")

# indicate directory with landscape outputs and relative paths for comparing various groups
path <- "outputs"
groups <- c("1k", "2k")

# image (width) resolution for outputted figures
resolution <- 1500    # in pixels
```

#### Generating average persistence landscapes

```{r}
# makes an average PL for each directory in "groups"
make_avgPL(path, groups)
```

#### Plotting average persistence landscapes

```{r}
# specify the number of landscape functions to plot
n <- 50    # this will plot the first "n" functions

# generate plots for average PLs
plot_avgPL(path, groups, n, resolution)
```

#### Plot difference(s) between averages landscapes

```{r}
# specify the number of landscape functions to plot
n <- 50    # this will plot the first "n" functions

# generate plots comparing pairs  (all combinations within "groups") from the specified groups
plot_avgPL_diff(path, groups, n, resolution)
```

#### Permutation test between groups

```{r}
# specify the max number of landscapes to use and the number of permutations to run
depth <- 30
reps <- 1000

# run test comparing pairs (all combinations within "groups") from the specified groups
perm_test(path, groups, depth, reps)
```
