# Discretization

------------------------------------------------------------------------

#### Dependencies

*Install the following libraries used for displaying discretized cell type images before running the pipeline.*

```{r}
install.packages(c("ggplot2", "ggforce", "cowplot"))
```

------------------------------------------------------------------------

## Functions and paths

*Pull generalized functions for processing the segmentation files.*

```{r}
# get helper functions
source("utilities.R")

# specify location path of segmented microscopy CSVs
csv_path <- "~/Research/All/Nanog_Gata6/1"
```

## Normalize signals with nuclear signal

*Based on cell nuclear stain, normalize other signal intensities to account for cell size, z-location, etc.*

```{r}
# recursively iterates through all segmentation CSV files in a directory
normalize(csv_path)
```

## Discretize channel intensities through generated thresholds

*First generate thresholds based on baseline data then discretize cell expression into cell types.*

```{r}
# indicate what percentile to generate threshold from (Ex. 75)
percentile <- 75

# specify directory paths of baseline CSVs for each threshold (one per channel)
gfp_thresh_path <- "~/Research/All/Nanog_Gata6/1/25"
rfp_thresh_path <- "~/Research/All/Nanog_Gata6/1/0"

# generate respective channel thresholds (1, ..., n) for discretizing cell intensities (here two are used)
paths <- c(gfp_thresh_path, rfp_thresh_path)    # in order as listed in segmentation CSV
thresholds <- get_thresholds(percentile, paths)

# -------------------------------------------------------------------------------------------

# specify directory paths for new (percentile-specific) discretized CSVs
disc_path <- "~/Research/Discretized/75/Nanog_Gata6/1"

# generate new CSVs with discretized expression intensities
discretize(csv_path, disc_path, thresholds)
```

## Generate images displaying cell types

*Using the cell types from the discretized data, generate* pseudo-microscopy images.

```{r}
# cell type image output parameters (in pixels)
size <- 3    # cell radius
outline <- 1   # cell outline (in black for overlap)
resolution <- 1500    # width of image

# CSV for specifying cell colors based on discretized cell type (Boolean string)
colors_path <- "cell_colors.csv"

# specify path location of discretized CSVs and image output
disc_path <- "~/Research/Discretized/75/Nanog_Gata6/1"
image_path <- "~/Research/Disc-Images/75/Nanog_Gata6/1"

# generate images from discretized cell types
generate_images(disc_path, image_path, colors_path, size, outline, resolution)
```
