---
title: "Topological data analysis"
output:
  html_document:
   df_print: paged
---
#### Dependencies
*Install the following libraries before running the TDA pipeline.*

```{r}
install.packages(c("TDA", "Matrix"))
```
---------------------------

### Load libraries

```{r libraries}
library("Matrix") #package for sparse matrices
library("TDA")
```

### Set working directory

```{r directory}
working_dir_path <- setwd("~/Documents/GitHub/TDA-Microscopy-Pipeline/TDA/general-pipeline")
```

### Load necessary functions

```{r functions}
source("utilities.R")
```

### Create a folder with a subfolder for each data group for storing saved tda computations
*You only need to run this block once*

```{r folders}
dir.create("saved_computations", showWarnings = FALSE)
saved_computations <- concat_path(working_dir_path, "saved_computations")

#choose names for two data groups
groups <- list("group_0", "group_25")

for (i in 1:length(groups)){
  dir.create(concat_path(saved_computations,groups[i]),  showWarnings = FALSE)
}
```

### Create a single csv file for storing necessary parameters (max birth and max death) obtained from all data groups
*You only need to run this block once*

```{r parameters-file}
max_values <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(max_values) <- c("maxbirth", "maxdeath")  

write.csv(max_values, concat_path(saved_computations,"max-values.csv"), row.names=FALSE)
```

### Choose cell types

```{r cell-types}
#specify location path of the file with the cell type codes and their colors
cell_types_path <- "~/Documents/GitHub/TDA-Microscopy-Pipeline/Discretization/cell_colors.csv"

#read all cell type codes from your designated file
cell_types_file <- read.csv(cell_types_path)
all_cell_types <- cell_types_file$Code

# indicate cell types for only including specific cells 
cell_types <- all_cell_types[2]   # this can either be a single value or a vector

index <- NaN    # default index for cell types is the last column

sprintf("Cell types used: %s", paste(cell_types, collapse=", "))
```

## I. Choose a data group and its location

```{r data-groups}
#choose a group folder where to save computations
group <- "group_0"

# specify location path of the csv files of that group
csv_dir_path <- "~/Documents/GitHub/TDA-Microscopy-Pipeline/Discretized-images/0"
```

## II. Compute persistence diagrams and save them
*We use Delaunay complex filtration to compute persistence diagrams*

```{r PDs}
compute_PDs(csv_dir_path, group, saved_computations, cell_types, index)
```

## III. Determine the max birth and max death among all data files
*This is needed to generate plots in both data groups with the same scale*

```{r parameters}
#If you want to generate plots in both data groups with the same scale, then before running this block, compute persistence diagrams for all data groups (i.e choose another data group and specify its location in I-block, and compute its persistence diagrams in II-block) 

max_values <- read.csv(concat_path(saved_computations, "max-values.csv"))

max_birth <- max(max_values$maxbirth)
max_death <- max(max_values$maxdeath)
```

### Plot persistence diagrams

```{r plot-PDs}
save_plot <- TRUE #set FALSE if you don't want to save plots

plot_PDs(group, max_birth, max_death, saved_computations, save_plot)
```

### Plot representative cycles that persist (live) over a certain threshold

```{r plot-cycles}
#choose a persistence threshold
persist_threshold <- 20 

save_plot <- TRUE #set FALSE if you don't want to save plots

#plot representative cycles with persistence above the chosen threshold
plot_representative_cycles(csv_dir_path, cell_types, persist_threshold, group, saved_computations, save_plot)
```

## IV. Compute persistence landscapes and save them
*You need to have persistence diagrams computed first*

```{r PLs}
#choose a discretization step (use the same for all groups)
discr_step <- 0.3

save_computations_csv <- TRUE #set TRUE if additionally to saving persistence landscapes in RData format you want to save each of them in a separate csv file. This option is provided if, for example, you plan to use persistence landscapes for further analysis in another programming language. Note that in a csv file row k corresponds to a persistence landscape function at depth k, where depth 1 corresponds to a outermost persistence landscape function, and column names correspond to x-values. 

compute_PLs(group, max_birth, max_death, discr_step, saved_computations, save_computations_csv)
```

### Plot persistence landscapes 

```{r plot-PLs}
save_plot <- TRUE #set FALSE if you don't want to save plots

plot_PLs(group, max_birth, max_death, discr_step, saved_computations, save_plot)
```

## V. Compute the average persistence landscape and save it
*You need to have persistence landscapes computed first*

```{r average-PLs}
compute_avgPL(group, max_birth, max_death, discr_step, saved_computations)
```

### Plot the average persistence landscape

```{r plot-average-PLs}
save_plot <- TRUE #set FALSE if you don't want to save plots

plot_avgPL(group, max_birth, max_death, discr_step, saved_computations, save_plot)
```

## At this point, you need to have computed persistence diagrams, persistence landscapes, and average persistence landscapes for BOTH groups

-----------------

## VI. Plot the difference of average persistence landscapes of two data groups
*You need to have computed average persistence landscapes for both groups*

```{r difference-plot}
save_plot <- TRUE #set FALSE if you don't want to save plots

plot_avgPLs_difference(groups, max_birth, max_death, discr_step, saved_computations, save_plot)
```

## VII. Perform a permutation test on persistence landscapes from two data groups
*You need to have computed persistence landscapes for both groups*

```{r permutation-test}
#number of permutations
num_repeats <- 10000

permutation_test_for_PLs(groups, saved_computations, num_repeats)
```

