# TDA-Microscopy-Pipeline

## Install libraries

The following libraries help generate the discretized images.

```{r}
install.packages(c("ggplot2", "ggforce", "cowplot"))
```

## Normalize GFP/RFP expression

This code normalizes the expression values for GFP/RFP based on the cell specific DAPI nuclear stain expression.

```{r}
# pull functions from functions.R script
source("functions.R")

# set working directory to location of microscopy CSVs
setwd("~/Documents/Research/TDA/R/TDA-Microscopy-Pipeline")

# iterate over treatment groups
for (sig in c("Nanog_Gata6", "Nanog_HA")) {
  for (conc in c(0, 5, 15, 25)) {
    for (well in c(1, 2, 3)) {
      for (loc in c(1, 2, 3, 4, 5)) {
        # get path to microscopy CSV file
        csv_path <- paste0("CSVs/",sig,"/",conc,"/",sig,"_",conc,"_",well,"_",loc,"/")
        csv_name <- "intensities.csv"

        # make new directory for storing normalized GFP/RFP values
        norm_path <- paste0("Normalized/",sig,"/",conc,"/")
        dir.create(norm_path, recursive=TRUE, showWarnings=FALSE)

        # get the new CSV filename and full path
        norm_name <- paste0(sig,"_",conc,"_",well,"_",loc,".csv")

        # read the CSV, normalize the GFP/RFP, and save to new CSV
        df <- read.csv(paste0(csv_path, csv_name))
        df <- normalize(df)
        save_to_csv(df, paste0(norm_path, norm_name))
      }
    }
  }
}
```

## Generate discretized CSVs and images

Thresholds will be generated for each percentile and then applied to the data.

```{r}
# pull functions from functions.R script
source("functions.R")

# set working directory to location of microscopy CSVs
setwd("~/Documents/Research/TDA/R/TDA-Microscopy-Pipeline/")

# hold actual values for dox concentrations and threshold percentiles
dox <- c(0, 5, 15, 25)
percs <- c(75, 80, 85, 90, 95)

# store distributions of cell types in a 2D list of 2x2 matrices
M <- list()
for (perc_index in c(1, 2, 3, 4, 5)) {
  M[[perc_index]] <- list()
  for (dox_index in c(1, 2, 3, 4)) {
    M[[perc_index]][[dox_index]] <- matrix(rep(0,4), ncol=2)
  }
}

# iterate over treatment groups
for (sig in c("Nanog_Gata6", "Nanog_HA")) {
  for (loc in c(1, 2, 3, 4, 5)) {
    # store location specific expression values for generating thresholds
    gfp_values <- list()
    rfp_values <- list()
    
    # load baseline CSVs for generating thresholds
    for (well in c(1, 2, 3)) {
      # directory path for baseline expression values
      dir_gfp <- paste0("Normalized/", sig, "/25/")
      dir_rfp <- paste0("Normalized/", sig, "/0/")
      
      # get CSV file names
      gfp_file <- paste0(sig, "_25_", well, "_", loc, ".csv")
      rfp_file <- paste0(sig, "_0_", well, "_", loc, ".csv")
      
      # read the CSV to data frames
      gfp_df <- read.csv(paste0(dir_gfp, gfp_file))
      rfp_df <- read.csv(paste0(dir_rfp, rfp_file))
      
      # hold location specific expression values for GFP/RFP
      gfp_values[[well]] <- gfp_df[, 6]
      rfp_values[[well]] <- rfp_df[, 7]
    }
    
    for (perc_index in c(1, 2, 3, 4, 5)) {
      # get the percentile value
      perc <- percs[[perc_index]]
      
      # generate thresholds based percentile
      gfp_threshold <- gen_threshold(perc, gfp_values)
      rfp_threshold <- gen_threshold(perc, rfp_values)
      
      for (dox_index in c(1, 2, 3, 4)) {
        # get dox concentration value
        conc <- dox[[dox_index]]
          
        for (well in c(1, 2, 3)) {
          # get path to CSV file and file name
          csv_path <- paste0("Normalized/", sig, "/", conc, "/")
          csv_name <- paste0(sig, "_", conc, "_", well, "_", loc, ".csv")
          
          # read the CSV, discretize the GFP/RFP values, and assign colors
          df <- read.csv(paste0(csv_path, csv_name))
          df <- discretize(df, gfp_threshold, rfp_threshold)
          df <- assign_color(df, show_double_low=TRUE)
          
          # get/make directory for storing discretized images and CSVs
          disc_path_im <- paste0("Discretized_Images/", perc, "/", sig, "/", conc, "/")
          disc_path_csv <- paste0("Discretized_Values/", perc, "/", sig, "/", conc, "/")
          dir.create(disc_path_csv, recursive=TRUE, showWarnings=FALSE)
          dir.create(disc_path_im, recursive=TRUE, showWarnings=FALSE)
      
          # get image and CSV names and save
          image_name <- paste0(sig, "_", conc, "_", well, "_", loc, ".png")
          csv_name <- paste0(sig, "_", conc, "_", well, "_", loc, ".csv")
          save_image(df, paste0(disc_path_im, image_name))
          save_to_csv(df, paste0(disc_path_csv, csv_name))
          
          # add to distribution table
          value <- gen_table(df, gfp_threshold, rfp_threshold)
          M[[perc_index]][[dox_index]] <- M[[perc_index]][[dox_index]] + value
        }
      }
    }
  }
}
```

## Show cell type distributions

Matrices containing the distribution of cell types based on the dox treatment and threshold percentile can be printed with the following code.

```{r}
# iterate over treatment groups
for (perc_index in c(1, 2, 3, 4, 5)) {
  for (dox_index in c(1, 2, 3, 4)) {
    # print a header based on the dox concentration and the threshold percentile
    cat("dox: ", dox[[dox_index]], ", percentile: ", percs[[perc_index]], "\n", sep="")
    
    # print the cell type distribution matrix
    print(M[[perc_index]][[dox_index]] / sum(M[[perc_index]][[dox_index]]))
    cat("\n")
  }
}
```
